---
title: "Vis II"
author: "Xiaoni Xu"
date: "2024-10-01"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)

```

# Exploratory analysis using data summaries
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10,
    month = lubridate::floor_date(date, unit = "month")) |>
  select(name, id, everything())
```

## Initial numeric explorations

Making the histogram
```{r}
weather_df |> 
  ggplot(aes(x = prcp)) + 
  geom_histogram()

weather_df |> 
  filter(prcp >= 1000)
```

```{r}
weather_df |> 
  filter(tmax >= 20, tmax <= 30) |> 
  ggplot(aes(x = tmin, y = tmax, color = name, shape = name)) + 
  geom_point(alpha = .75)
```

## `group_by`
```{r}
weather_df |>
  group_by(name, month)
```

## Counting things
```{r}
weather_df |>
  group_by(month) |>
  summarize(n_obs = n())
```

group by more than one variable
```{r}
weather_df |>
  group_by(name, month) |>
  summarize(n_obs = n())
```
In both cases, the result is a dataframe that includes the grouping variable(s) and the desired summary.

To count things, you could use `count`() in place of `group_by`() and `summarize`() if you remember that this function exists. I’ll also make use of the `name` argument in `count`, which defaults to "n".
```{r}
weather_df |>
  count(month, name = "n_obs")
```

You can use `summarize`() to compute multiple summaries within each group. As an example, we count the number of observations in each month and the number of distinct values of `date` in each month.
```{r}
weather_df |>
  group_by(month) |>
  summarize(
    n_obs = n(),
    n_days = n_distinct(date))
```

## 2x2

```{r}
weather_df %>% 
  drop_na(tmax) %>% 
  filter(name != "Molokai_HI") %>% 
  mutate(cold = case_when(
    tmax < 5 ~ "cold",
    tmax >= 5 ~ "not_cold"
  )) %>% 
  janitor::tabyl(name, cold)
# don't use the `table` function
```

## general numeric summaries

let's try some other useful summaries

summarize then plot...

```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE),
    median_tmin = median(tmin, na.rm = TRUE),
    sd_prcp = sd(prcp, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = month, y = mean_tmax, color = name)) +
  geom_point() +
  geom_line()
```

format for readers
```{r}
weather_df %>% 
  group_by(name, month) %>% 
  summarize(
    mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>%
  pivot_wider(
    names_from = name, 
    values_from = mean_tmax
  ) %>% 
  knitr::kable(
  digits = 3,
  col.names = c("Month", "Central Park", "Molokai", "Waterhole"))
  
```

