---
title: "Vis II"
author: "Xiaoni Xu"
date: "2024-10-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)
library(viridis)
library(haven)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

As a starting point, let’s revisit the scatterplot of tmax against tmin made in Visualization Pt 1.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3)
```

## Labels

There are a variety of useful ways to change the appearance of your plot, especially if your graphic is intended to be viewed by others. One of the most important things you can do is provide informative axis labels, plot titles, and captions, all of which can be controlled using labs().

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package"
  )
```

## Scales

Aesthetic mappings determine which variables map to which plot attributes. These mappings have reasonable default behaviors, but can be modified through scales.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15º C", "0", "15"))
```

There are a variety of other `scale_x_*` and `scale_y_*` options – it can be helpful to know how and where these are controlled.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 20), 
    labels = c("-15ºC", "0", "20"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```

For example, arguments to `scale_color_hue()` control the color scale and the name in the plot legend.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(h = c(100, 300))
```

Trying to create your own color scheme usually doesn’t go well; use the `viridis` package instead. There are several options, but the default color scheme works nicely!


```{r}
ggp_temp_plot = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )

ggp_temp_plot
```

## Themes
```{r}
ggp_temp_plot + 
  theme(legend.position = "bottom")
```
`legend.position = "none"` will remove the legend. This is helpful when multiple plots use the same color scheme or when the legend is obnoxious for some other reason.



```{r}
ggp_temp_plot + 
  theme_bw() + 
  theme(legend.position = "bottom")
```

```{r}
ggp_temp_plot + 
  theme_classic() + 
  theme(legend.position = "bottom")
```

```{r}
ggp_temp_plot + 
  ggthemes::theme_excel() + 
  theme(legend.position = "bottom")
```


Revisit the plot showing tmax against date for each location. Use labels, scale options, and theme changes to improve the readability of this plot.
```{r}
weather_df %>%
  ggplot(aes(x = date, y = tmax, color = name, SIZE = prcp)) +
  geom_point(size = 2, alpha = 0.3) +  # Adjust point size and transparency for readability
  geom_smooth(se = FALSE) +
  scale_color_viridis(
    name = "Location", 
    discrete = TRUE, 
    option = "viridis"  # Use the 'mako' color palette
  ) +
  labs(
    title = "Date-Temperature Plot",
    x = "Date",
    y = "Maximum Daily Temperature (°C)",
    color = "Location",
    caption = "Maximum temperature against date for each location",
    size = "Precipitation"
  ) +
  theme_minimal()+
  theme(
    plot.title = element_text(hjust = 0.5),  # Center the title
    axis.text.x = element_text(angle = 45, hjust = 1)  # Rotate x-axis labels for better fit
  )
  
```

Extra bonus stuff in `ggplot`

Use different datasets in different `geom`s
```{r}
central_park_df =
  weather_df %>% 
  filter(name == "CentralPark_NY")

molokai_df = 
  weather_df %>% 
  filter(name == "Molokai_HI")

molokai_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```


## Multiple panels

```{r}
weather_df %>% 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density() +
  facet_grid(. ~ name)
```

```{r}
tmax_tmin_p = 
  weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) + 
  geom_point(alpha = .5) +
  theme(legend.position = "none")

prcp_dens_p = 
  weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_density(alpha = .5) + 
  theme(legend.position = "none")

tmax_date_p = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  theme(legend.position = "bottom")

(tmax_tmin_p + prcp_dens_p) / tmax_date_p
```

## Data Manipulation

You can change the order level of a factor variable to your specified preference using `forcats::fct_relevel` or according to the value of another variable using `forcats::fct_reorder`.

A first example reorders `name` “by hand”:
```{r}
weather_df |>
  mutate(name = forcats::fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), color = "blue", alpha = .5) + 
  theme(legend.position = "bottom")
```

A more difficult situation relates to data tidyiness. Suppose I wanted to create a three-panel plot showing densities for `tmax` and `tmin` within each location. More concretely, I want to be able to facet panels across the `name` variable, and create separate densities for `tmax` and ``tmin` in each panel. Unfortunately, `weather_df` isn’t organized in a way that makes this easy.

One solution would recognize that `tmax` and `tmin` are separate observation types of a shared `temperature` variable. With this understanding, it’s possible to tidy the `weather_df` and make the plot directly:

```{r}
weather_df |>
  select(name, tmax, tmin) |> 
  pivot_longer(
    tmax:tmin,
    names_to = "observation", 
    values_to = "temp") |> 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) + 
  facet_grid(~name) + 
  viridis::scale_fill_viridis(discrete = TRUE)
```

The code below imports and tidies the PULSE data, and creates a plot showing BDI score across visits. Some steps that are helpful in retrospect are using `pivot_longer` to organize the BDI score and visit time variables, and organizing the visit time variable into a factor with an informative ordering.

```{r}
pulse_data = 
  haven::read_sas("./data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer(
    bdi_score_bl:bdi_score_12m,
    names_to = "visit", 
    names_prefix = "bdi_score_",
    values_to = "bdi") |>
  select(id, visit, everything()) |>
  mutate(
    visit = recode(visit, "bl" = "00m"),
    visit = factor(visit, levels = str_c(c("00", "01", "06", "12"), "m"))) |>
  arrange(id, visit)

ggplot(pulse_data, aes(x = visit, y = bdi)) + 
  geom_boxplot()
```

Revisit FAS data
```{r}
pup_data = 
  read_csv("./data/FAS_pups.csv", na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  mutate(
    sex = 
      case_match(
        sex, 
        1 ~ "male", 
        2 ~ "female"))

litter_data = 
  read_csv("./data/FAS_litters.csv", na = c("NA", ".", "")) |>
  janitor::clean_names() |>
  separate(group, into = c("dose", "day_of_tx"), sep = 3)

fas_data = left_join(pup_data, litter_data, by = "litter_number") 

fas_data |> 
  select(sex, dose, day_of_tx, pd_ears:pd_walk) |> 
  pivot_longer(
    pd_ears:pd_walk,
    names_to = "outcome", 
    values_to = "pn_day") |> 
  drop_na() |> 
  mutate(outcome = forcats::fct_reorder(outcome, pn_day, median)) |> 
  ggplot(aes(x = dose, y = pn_day)) + 
  geom_boxplot() + 
  facet_grid(day_of_tx ~ outcome)
```






