---
title: "Vis II"
author: "Xiaoni Xu"
date: "2024-10-01"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(tidyverse)
library(patchwork)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

weather_df
```

As a starting point, let’s revisit the scatterplot of tmax against tmin made in Visualization Pt 1.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3)
```

## Labels

There are a variety of useful ways to change the appearance of your plot, especially if your graphic is intended to be viewed by others. One of the most important things you can do is provide informative axis labels, plot titles, and captions, all of which can be controlled using labs().

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .3) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package"
  )
```

## Scales

Aesthetic mappings determine which variables map to which plot attributes. These mappings have reasonable default behaviors, but can be modified through scales.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15º C", "0", "15"))
```

There are a variety of other `scale_x_*` and `scale_y_*` options – it can be helpful to know how and where these are controlled, although I usually would have to google how to do what I want.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_x_continuous(
    breaks = c(-15, 0, 15), 
    labels = c("-15ºC", "0", "15"),
    limits = c(-20, 30)) + 
  scale_y_continuous(
    trans = "sqrt", 
    position = "right")
```

For example, arguments to scale_color_hue() control the color scale and the name in the plot legend.

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    color = "Location",
    caption = "Data from the rnoaa package") + 
  scale_color_hue(h = c(100, 300))
```





