---
title: "Vis I"
author: "Xiaoni Xu"
date: "2024-09-26"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)
```

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = case_match(
      id, 
      "USW00094728" ~ "CentralPark_NY", 
      "USW00022534" ~ "Molokai_HI",
      "USS0023B17S" ~ "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

Making the first plot

```{r}
ggplot(weather_df, 
       aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
ggp_weather_scatterplot = 
  weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point()

ggp_weather_scatterplot
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3, size = .8) + 
  geom_smooth(se = FALSE)
```

Where you definte aesthetics can matter

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(aes(color = name), alpha = .3, size = .8) + 
  geom_smooth(se = FALSE)
```

Use faceting
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) 

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point(alpha = .3) +
  geom_smooth(se = FALSE) +
  facet_grid(name ~ .) # horizontal
```

Let's make a somewhat more interesting scatterplot

```{r}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name, size = prcp)) +
  geom_point(alpha = .3) + #see the seasonal trends
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

Learning assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).
```{r}
weather_df %>%
  filter(name == "CentralPark_NY") %>%  # Filter for Central Park
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32
  ) %>% 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point() +
  geom_smooth(method = 'lm', se = FALSE)
```

Smol things
```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_hex()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_point(color = "blue")

weather_df %>% 
  ggplot(aes(x = tmin, y = tmax), color = "blue") + # "blue" is not in the dataset
  geom_point()
```

## Univariate plots

```{r}
weather_df %>% 
  ggplot(aes(x = tmin)) +
  geom_histogram()
```

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) + # difficult to understand and hard to read if use color = name, according to JEFF
  geom_histogram(position = "dodge")
```

How would JEFF fix this?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_histogram() +
  facet_grid(. ~ name)
```

Maybe a density plot?

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, fill = name)) +
  geom_density(alpha = .3)
```

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_boxplot()
```

Violin plots

```{r}
weather_df %>% 
  ggplot(aes(x = name, y = tmin, fill = name)) +
  geom_violin()
```

Ridge plot

```{r}
weather_df %>% 
  ggplot(aes(x = tmin, y = name)) +
  geom_density_ridges()
```

Learning Assessment: Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.

```{r}
# Histogram
histogram_plot <- weather_df %>%
  ggplot(aes(x = prcp, fill = name)) +
  geom_histogram(position = "dodge", bins = 30, alpha = 0.7) +
  labs(title = "Precipitation Histogram") +
  theme_minimal()

# Density Plot
density_plot <- weather_df %>%
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = 0.5) +
  labs(title = "Precipitation Density Plot") +
  theme_minimal()

# Boxplot
boxplot_plot <- weather_df %>%
  ggplot(aes(x = name, y = prcp, fill = name)) +
  geom_boxplot() +
  labs(title = "Precipitation Boxplot") +
  theme_minimal()

# Violin Plot
violin_plot <- weather_df %>%
  ggplot(aes(x = name, y = prcp, fill = name)) +
  geom_violin() +
  labs(title = "Precipitation Violin Plot") +
  theme_minimal()

# Ridge Plot
ridge_plot <- weather_df %>%
  ggplot(aes(x = prcp, y = name, fill = name)) +
  geom_density_ridges(aes(height = ..density..), alpha = 0.7) +  # Ensure height aesthetic is specified
  labs(title = "Precipitation Ridge Plot") +
  theme_minimal()

# Combine all plots
combined_plot <- (histogram_plot | density_plot) / 
                 (boxplot_plot | violin_plot) / 
                 ridge_plot

# Display the combined plot
print(combined_plot)

weather_df %>%
  filter(prcp > 10, prcp < 1000) %>% 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .3)
```

## Saving and embedding plots

```{r}
ggp_weather =
  weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()

# create a folder to save all the stuff if you know you need to make a lot of plots
ggsave("plots/ggp_weather.pdf", ggp_weather, width = 8, height = 6)
```

embedding plots
```{r, fig.width = 6, fig.asp = .6}
weather_df %>% 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
```



